\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-04-07 forwards-and-backwards.tex
notes from meeting

@@

\begin{lemma}
    \label{lem:conditions on r}
If either
\begin{enumerate}
\item
\label{control through r} 
    $|\partial_{x_i}r(x,\smooth{r}\eta(x))|$ and 
    $|\partial_{x_ix_j}r(x,\smooth{r}\eta(x))|$
    are uniformly bounded for $x\in\IR^d$, $\eta\in\measures$;
\item 
\label{control through gamma}
    or, $m^2\gamma(x,m)$ is uniformly bounded and there exists $C<\infty$
    such that for $\theta$ sufficiently large, and all $x\in\IR^d$, $\eta\in\measures$,
    \[
    \theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)q_\theta(x,dy)\leq C\smooth{\gamma}\eta(x),
    \]
    and
    \[\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)^2q_\theta(x,dy)
    \leq C(\smooth{\gamma}\eta(x))^2.
    \]
\end{enumerate} 
then
$|\gamma(x, \smooth{\gamma} \eta(x)) \theta \int (r(y, \smooth{r} \eta(y)) - r(x, \smooth{r} \eta(x))) q_\theta(x, dy)|$
is uniformly bounded for $x \in \IR^d$ and $\eta \in \measures$.
\end{lemma}

@@

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex1a/fkpp_123.locations}
        \includegraphics{figures/ex1b/fkpp_123.locations}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        Blue curves are contours from a kernel density estimate.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover a superprocess with nonlinear death rates corresponding to logistic growth
that is a continuous limit of the Bolker-Pacala model \citep{bolker/pacala:1997,bolker/pacala:1999}.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson--Girsanov transform tells us that we have uniqueness in this special 
case of the superprocess with nonlinear death rates, and 
Perkins stochastic calculus (and its adaptation to a 
lookdown setting) provides uniqueness for cases with interactions in the 
dispersal mechanism of
the superprocess. We refer 
to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that the martingale problem is well-posed in the case $\alpha>0$.

@@

Can these two situations be distinguished from summaries of lineage movement?
The first has lineage generator
\[
    f \mapsto \Lgen f = r \gamma \left( \Delta f + 2 \grad \log(\gamma w) \cdot \grad f \right),
\]
while the second has lineage generator $f \mapsto \lambda(x) \Lgen f(x)$.
In other words,
although the stationary profile of the population is unchanged when we scale
local establishment and death by $\lambda$,
the motion of lineages is sped up locally by $\lambda$.
This corresponds to making areas with $\lambda > 1$ more ``sink-like'' 
and $\lambda < 1$ more ``source-like'':
if $\lambda(x) > 1$, then at $x$ both the death rate and probability of 
establishment of new individuals are higher.
As a result, lineages in the second model spend more time in areas with $\lambda < 1$,
i.e., those areas have higher reproductive value, %long-term fitness
something that is, in principle, discernible from genetic data.

@@

The lookdown processes have been carefully constructed so that
observations about the past spatial positions of individuals in 
the population do not give us any information about the assignment of 
individuals to levels.
In other words, the dynamics of the lookdown
process preserve the conditionally uniform (or in the limit, conditionally Poisson) structure
-- if started with uniform levels,
levels are uniform at all future times.
Moreover, if we average over levels in the expression for the generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we recover the generator for the population process.
%(visible in equation~\eqref{eqn:eta_martingale}
%or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (along with some boundedness conditions)
the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2019}
tells us that by ``removing labels'' from the lookdown process $\lp$
we recover the population process $\eta$.
% In particular, for any
% solution to the martingale problem of Definition~\ref{defn:mgale_construction}
% % of equations~\eqref{eqn:eta_martingale} and \eqref{eqn:prelimit_martingale_variation}
% there is a % lookdown representation $(\lp^N_t)_{t \ge 0}$
% solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
% for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])/N$
% and such that at each $t$, conditioned on $(\eta^N_s)_{s \le t}$, the levels of $\lp_t^N$
% are independently and uniformly distributed on $[0, N]$.
% Similarly, in the scaling limit:
% for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of 
% Theorem~\ref{nonlocal_convergence}
% %there is a lookdown process $(\lp^N_t)_{t \ge 0}$
% there is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
% for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
% and such that at each $t$, conditional on $(\eta_s)_{s\leq t}$, 
% $\lp_t$ is conditionally Poisson with mean measure $\eta_t(dx) \times du$.
Furthermore, convergence of the lookdown processes would imply convergence of the population processes
\citep[Thm A.9 in][]{kurtz/rodrigues:2011}.

@@

\begin{theorem}
    \label{thm:lookdown_convergence}
    Let $(\lp_t^N)$ satisfy Definition~\ref{defn:lookdown_mgale}
    and assume that as $N \to \infty$,
    $\theta \to \infty$ in such a way that $\theta/N \to \alpha$.
    Let $\eta_0^N = \kappa(\lp_0^N)$ and
    suppose also that $\eta^N_0 \to \eta_0$ (as finite measures on $\IR^d$),
    and that for each $N$, $\lp^N_0$ is conditionally uniform given $\eta_0^N$
    in the sense of \eqref{eqn:conditionally_uniform}.
    Then, any sequence of stochastic processes $(\lp_t^N)_{t \ge 0}$
    has a subsequence which converges in distribution as $N \to \infty$
    to a measure-valued process $(\lp_t)_{t \ge 0}$
    with $\lp_0$ conditionally Poisson given $\eta_0 = \kappa(\lp_0)$
    in the sense of \eqref{eqn:conditionally_poisson},
    that is a solution to the martingale problem of Definition~\ref{defn:limiting_lookdown_mgale}.
\end{theorem}

@@

In the following sections we present formal proofs of our three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain a limit in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
In Section~\ref{subsec:one step convergence proof}
we prove our third scaling limit, in which 
we pass from our population model 
directly to a reaction diffusion equation without going via a nonlocal limit.
Finally, in Section~\ref{sec:lookdown_proofs} we prove our results concerning both convergence of the lookdown representation
and characterization of the lineage process in the deterministic case.

@@

Below we will have frequent use for the quantity
\begin{equation}
\label{generator Bf}
    B_f(x, \eta) = \theta \int_{\IR^d} (f(y) r(y, \eta) - f(x) r(x, \eta)) q_\theta(x, dy) .
\end{equation}

@@

Now suppose that instead
Condition~\ref{control through gamma} of Lemma~\ref{lem:conditions on r} is satisfied.
is satisfied.
First note that
\begin{multline*}
    r(x, \smooth{r}\eta(y)) - r(x, \smooth{r}\eta(x))
    \\
=
    (\smooth{r}\eta(y) - \smooth{r}\eta(x)) r'(x,\smooth{r}\eta(x))
    + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 r''(x, \overline{m}),
\end{multline*}
where $\overline{m}= \kappa'\smooth{r}\eta(x) +(1-\kappa') 
\smooth{r}\eta(y)$ for some $0\leq \kappa'\leq 1$, and
we have used $r'$ and $r''$ to denote the first and second derivatives of $r(x,m)$
with respect to the second argument.
So, in this case
\begin{align*}
&
\left| \theta 
    \int_{\IR^n} 
        \big\{
            r\big(x,\smooth{r}\eta(y)\big)
            -
            r\big(x,\smooth{r}\eta(x)\big)
        \big\}
    q_{\theta}(x,dy) 
\right| 
\\ &\qquad
\le
   \|r'\|_\infty
    \left|
        \theta \int_{\IR^d}
            (\smooth{r}\eta(y) - \smooth{r}\eta(x))
        q_\theta(x, dy)
    \right|
    +
   \|r''\|_\infty
    \left|
        \theta \int_{\IR^d}
            (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
        q_\theta(x, dy)
    \right| .
\end{align*}
Therefore,
under Condition~\ref{control through gamma}, $|B_1(x,\eta)|$ is bounded by 
a constant times $\smooth{\gamma}\eta(x)+(\smooth{\gamma}\eta(x))^2$;
since $m^2\gamma(m)$ is bounded, so is $|\gamma(x,\eta) B_1(x,\eta)$.
\end{proof}

@@

\begin{lemma}
\label{no mass at infty}
Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
let $(\eta_t)_{t\geq 0}$ be a limit point of $\{(\eta^N_t)_{t\geq 0}\}_{N\geq 1}$. 
For any $\delta>0$, 
\[
\IP\big[\eta_t\big(\{\|x\|>R\}\big)>\delta\big]\to 0\qquad\mbox{ as }R\to\infty.
\]
\end{lemma}
\begin{proof}[Sketch]
To avoid the idea being buried in notation, we sketch the proof in the 
special case of $d=1$ and $\DG=\Delta$.
For any $f\in C_b^\infty(\IR)$,
\begin{multline}
\label{special case first moment equation}
\IE[\langle f,\eta_t\rangle] 
=\IE[\langle f,\eta_0\rangle]
\\
+\IE\Big[\int_0^t\Big\langle\gamma(x, \smooth{\gamma}\eta_s(x))
\Delta \big(f(\cdot)r(\cdot,\smooth{r}\eta_s(\cdot))\big)(x)
+F(x,\smooth{F}\eta_s(x)) f(x),
\eta_s(dx)\Big\rangle ds\Big].
\end{multline} 
We take a sequence of test functions $f_n$ increasing to
$\exp(\lambda x)$, with $\lambda>0$. Having passed
to the limit,
substituted in~(\ref{special case first moment equation}) and 
expanded the term involving the Laplacian, we obtain
\begin{multline}
\label{exponential moment}
    \IE[\langle \exp(\lambda x),\eta_t\rangle]
    =
    \IE[\langle \exp(\lambda x),\eta_0\rangle]
    +
    \IE\Big[\int_0^t\Big\langle\gamma(x, \smooth{\gamma}\eta_s(x))
    \exp(\lambda x)
    \Big\{
        \lambda^2 r(x, \smooth{r}\eta_s(x))
    \\ {}
        + 2\lambda\partial_x\big(r(\cdot,\smooth{r}\eta_s(\cdot)\big)(x)
        + \partial_{xx} r(\cdot,\smooth{r}\eta_s(\cdot)\big)(x)
    \Big\} 
    + F(x,\smooth{F}\eta_s(x)),
    \eta_s(dx)\Big\rangle ds\Big] .
\end{multline} 
Evidently, in much the same way as in the proof of
Lemma~\ref{lem:eta_f_bound},
our conditions on $r$ and $F$ ensure
that the integrand on the right hand side 
of~(\ref{exponential moment})
is bounded by $C\langle\exp(\lambda x),\eta_s(dx)\rangle$ for some constant
$C$. Gronwall's inequality then implies that 
$\IE[\langle\exp(\lambda x),\eta_t(dx)\rangle]$ 
is uniformly bounded for $t\in [0,T]$, and an application of Markov's inequality
tells us that for any $\delta >0$,
$\IP[\eta_t([R,\infty])>\delta]\to 0$ as $R\to\infty$. A symmetric argument
yields the corresponding bound for $\IP[\eta_t([-\infty, -R])>\delta$ and the 
result is proved.
\end{proof}

@@

@@

For the second claim, first define $\tau(t) = T(t)-t$.
Since if $\tau(t) \ge -(\epsilon^2 + t)/2$,
then $1/(\epsilon^2 + T(t)) \le 2/(\epsilon^2 + t)$,
while $\epsilon^2 + T(t) \ge \epsilon^2$ always,
then partitioning over
$\{ \tau(t) \ge -(\epsilon^2 + t)/2 \}$ and its complement,
\begin{align}
   \| \psi^{\epsilon,x} \|_\infty
    &=
        \IE\left[ 
            \frac{1}{\big(2 \pi (\epsilon^2+T(t))\big)^{d/2}}
        \right] \nonumber
    \\ & \le
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        + 
        \frac{C}{\epsilon^d}
        \IP\left\{
            \tau(t) < - (\epsilon^2 + t) / 2
        \right\} \label{eqn:psi_infty_bound}
\end{align}
Now, observe that since $\IE[e^{-\Pi(\theta t)}] = \exp(-\theta t (1 - e^{-1}))$,
by Markov's inequality,
\begin{align}
    \IP\left\{
        \tau(t) < - \frac{\epsilon^2 + t}{2}
    \right\}
    &=
    \IP\left\{
        e^{-\Pi(\theta t)} > e^{-\theta (t - \epsilon^2) / 2}
    \right\}
\nonumber
\\
&\leq \frac{\IE[\exp\big(-\Pi(\theta t)\big)]}{\exp\big(-\theta(t-\epsilon^2)/2\big)}
\nonumber
    \\&=
    \frac{
        \exp(- \theta t(1 - e^{-1}))
    }{
        \exp(- \theta (t - \epsilon^2) / 2)
    }
\nonumber
    \\&=
    \exp\left\{ - \delta \theta t - \frac{\theta \epsilon^2}{2} \right\} ,
\label{bound for negative tau}
\end{align}
where $\delta = 1/2 - e^{-1} > 0$.
The second term in \eqref{eqn:psi_infty_bound} is therefore bounded by
\[
    C \left(1 + \frac{t}{\epsilon^2}\right)^{d/2} e^{-\delta \theta t - \epsilon^2 \theta / 2}
    \frac{1}{(\epsilon^2 + t)^{d/2}} .
\]
Now observe that the derivative (with respect to $t$) of 
$e^{-\delta \theta t - \epsilon^2 \theta / 2} (1 + t/\epsilon^2)^{d/2}$ 
is
\[
    \left( \frac{d}{2 \epsilon^2} - \left(1 + \frac{t}{\epsilon^2}\right) \delta \theta \right)
    \left( 1 + \frac{t}{\epsilon^2} \right)^{d/2 - 1} e^{-\delta \theta t - \epsilon^2 \theta / 2} ,
\]
which is negative if $\theta(\epsilon^2 + t) > d/2 \delta$.
At the maximum, $(\epsilon^2 + t) = d \delta / (2 \theta)$,
so $(1 + t/\epsilon^2) = d \delta/(2 \theta \epsilon^2)$,
and so this quantity is bounded uniformly over only $t$ but also $\epsilon$.
\comment{That's where we use bounded-below-ness of $\theta \epsilon^2$.}
Therefore, the second term in \eqref{eqn:psi_infty_bound} 
is bounded by $C / (\epsilon^2 + t)^{d/2}$
(for a different constant $C$).
Substituting this into \eqref{eqn:psi_infty_bound} yields the result.
\end{proof}

@@

\comment{Probably should put a bit more detail in here?}

>>>>

\comment{get equations to restart numbering in sections}

@@

\begin{lemma}
    \label{lem:conditions on r}
If either
    \comment{which assumptions from 2.7 are in force? since in (1) we need $\gamma$ to be bounded, which is condition 3 above}
\begin{enumerate}
\item
\label{control through r} 
    $|\partial_{x_i}r(x,\smooth{r}\eta(x))|$ and 
    $|\partial_{x_ix_j}r(x,\smooth{r}\eta(x))|$
    are uniformly bounded for $x\in\IR^d$, $\eta\in\measures$;
\item 
\label{control through gamma}
    or, $m^2\gamma(x,m)$ is uniformly bounded and there exists $C<\infty$
    such that for $\theta$ sufficiently large, and all $x\in\IR^d$, $\eta\in\measures$,
    \[
    \theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)q_\theta(x,dy)\leq C\smooth{\gamma}\eta(x),
    \]
    and
    \[\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)^2q_\theta(x,dy)
    \leq C(\smooth{\gamma}\eta(x))^2.
    \]
\end{enumerate} 
then
$|\gamma(x, \smooth{\gamma} \eta(x)) \theta \int (r(y, \smooth{r} \eta(y)) - r(x, \smooth{r} \eta(x))) q_\theta(x, dy)|$
is uniformly bounded for $x \in \IR^d$ and $\eta \in \measures$.
\end{lemma}

@@

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex1a/fkpp_123.locations}
        \includegraphics{figures/ex1b/fkpp_123.locations}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        Blue curves are contours from a kernel density estimate.
        \comment{remove blue curves}
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover a superprocess with nonlinear death rates corresponding to logistic growth
\comment{cite Etheridge 2004}
that is a continuous limit of the Bolker-Pacala model \citep{bolker/pacala:1997,bolker/pacala:1999}.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson--Girsanov transform tells us that we have uniqueness in this special 
case of the superprocess with nonlinear death rates, and 
Perkins stochastic calculus (and its adaptation to a 
lookdown setting) provides uniqueness for cases with interactions in the 
dispersal mechanism of
the superprocess. We refer 
to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that the martingale problem is well-posed in the case $\alpha>0$.

@@

Can these two situations be distinguished from summaries of lineage movement?
The first has lineage generator
\[
    f \mapsto \Lgen f = r \gamma \left( \Delta f + 2 \grad \log(\gamma w) \cdot \grad f \right),
\]
while the second has lineage generator $f \mapsto \lambda(x) \Lgen f(x)$.
In other words,
although the stationary profile of the population is unchanged when we scale
local establishment and death by $\lambda$,
the motion of lineages is sped up locally by $\lambda$.
This corresponds to making areas with $\lambda > 1$ more ``sink-like'' 
and $\lambda < 1$ more ``source-like'':
if $\lambda(x) > 1$, then at $x$ both the death rate and probability of 
establishment of new individuals are higher.
As a result, lineages in the second model spend more time in areas with $\lambda < 1$,
i.e., those areas have higher reproductive value, %long-term fitness
something that is, in principle, discernible from genetic data.
\comment{clarify why: smaller $N_e$}

@@

The lookdown processes have been carefully constructed so that
observations about the past spatial positions of individuals in 
the population do not give us any information about the assignment of 
individuals to levels.
In other words, the dynamics of the lookdown
process preserve the conditionally uniform (or in the limit, conditionally Poisson) structure
-- if started with uniform levels,
levels are uniform at all future times.
Moreover, if we average over levels in the expression for the generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we recover the generator for the population process.
%(visible in equation~\eqref{eqn:eta_martingale}
%or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (along with some boundedness conditions)
the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2019}
tells us that by ``removing labels'' from the lookdown process $\lp$
we recover the population process $\eta$.
% In particular, for any
% solution to the martingale problem of Definition~\ref{defn:mgale_construction}
% % of equations~\eqref{eqn:eta_martingale} and \eqref{eqn:prelimit_martingale_variation}
% there is a % lookdown representation $(\lp^N_t)_{t \ge 0}$
% solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
% for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])/N$
% and such that at each $t$, conditioned on $(\eta^N_s)_{s \le t}$, the levels of $\lp_t^N$
% are independently and uniformly distributed on $[0, N]$.
% Similarly, in the scaling limit:
% for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of 
% Theorem~\ref{nonlocal_convergence}
% %there is a lookdown process $(\lp^N_t)_{t \ge 0}$
% there is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
% for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
% and such that at each $t$, conditional on $(\eta_s)_{s\leq t}$, 
% $\lp_t$ is conditionally Poisson with mean measure $\eta_t(dx) \times du$.
% Furthermore, convergence of the lookdown processes would imply convergence of the population processes
% \citep[Thm A.9 in][]{kurtz/rodrigues:2011}.

@@

\begin{theorem}
    \label{thm:lookdown_convergence}
    Let $(\lp_t^N)$ satisfy Definition~\ref{defn:lookdown_mgale}
    and assume that as $N \to \infty$,
    $\theta \to \infty$ in such a way that $\theta/N \to \alpha$.
    Let $\eta_0^N = \kappa(\lp_0^N)$ and
    suppose also that $\eta^N_0 \to \eta_0$ in $\measures$,
    and that for each $N$, $\lp^N_0$ is conditionally uniform given $\eta_0^N$
    in the sense of \eqref{eqn:conditionally_uniform}.
    Then, any sequence of stochastic processes $(\lp_t^N)_{t \ge 0}$
    has a subsequence which converges in distribution as $N \to \infty$
    to a measure-valued process $(\lp_t)_{t \ge 0}$
    with $\lp_0$ conditionally Poisson given $\eta_0 = \kappa(\lp_0)$
    in the sense of \eqref{eqn:conditionally_poisson},
    that is a solution to the martingale problem of Definition~\ref{defn:limiting_lookdown_mgale}.
\end{theorem}

@@

\comment{this section (6) is no long just "proofs" since they're in section 7 and 8 too}
\comment{make sure this reads okay and describes just this section (6); the end-of-the-intro paragraph describes structure}
In the following sections we present formal proofs of our three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain a limit in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
In Section~\ref{subsec:one step convergence proof}
we prove our third scaling limit, in which 
we pass from our population model 
directly to a reaction diffusion equation without going via a nonlocal limit.
Finally, in Section~\ref{sec:lookdown_proofs} we prove our results concerning both convergence of the lookdown representation
and characterization of the lineage process in the deterministic case.

@@

Below we will have frequent use for the quantity
\comment{perhaps $B$ should have a superscript $\theta$}
\begin{equation}
\label{generator Bf}
    B_f(x, \eta) = \theta \int_{\IR^d} (f(y) r(y, \eta) - f(x) r(x, \eta)) q_\theta(x, dy) .
\end{equation}

@@

Now suppose that instead
Condition~\ref{control through gamma} of Lemma~\ref{lem:conditions on r} is satisfied.
is satisfied.
First note that
\begin{multline*}
    r(x, \smooth{r}\eta(y)) - r(x, \smooth{r}\eta(x))
    \\
=
    (\smooth{r}\eta(y) - \smooth{r}\eta(x)) r'(x,\smooth{r}\eta(x))
    + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 r''(x, \overline{m}),
\end{multline*}
where $\overline{m}= \kappa'\smooth{r}\eta(x) +(1-\kappa') 
\smooth{r}\eta(y)$ for some $0\leq \kappa'\leq 1$, and
we have used $r'$ and $r''$ to denote the first and second derivatives of $r(x,m)$
with respect to the second argument.
So, in this case
\begin{align*}
&
\left| \theta 
    \int_{\IR^n} 
        \big\{
            r\big(x,\smooth{r}\eta(y)\big)
            -
            r\big(x,\smooth{r}\eta(x)\big)
        \big\}
    q_{\theta}(x,dy) 
\right| 
\\ &\qquad
\le
   \|r'\|_\infty
    \left|
        \theta \int_{\IR^d}
            (\smooth{r}\eta(y) - \smooth{r}\eta(x))
        q_\theta(x, dy)
    \right|
    +
   \|r''\|_\infty
    \left|
        \theta \int_{\IR^d}
            (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
        q_\theta(x, dy)
    \right| .
\end{align*}
Therefore,
under Condition~\ref{control through gamma}, $|B_1(x,\eta)|$ is bounded by 
a constant times $\smooth{\gamma}\eta(x)+(\smooth{\gamma}\eta(x))^2$;
since $m^2\gamma(m)$ is bounded, so is $|\gamma(x,\eta) B_1(x,\eta)|$.
\end{proof}

@@

\begin{lemma}
\label{no mass at infty}
Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
let $(\eta_t)_{t\geq 0}$ be a limit point of $\{(\eta^N_t)_{t\geq 0}\}_{N\geq 1}$. 
For any $\delta>0$, 
\[
\IP\big[\eta_t\big(\{\|x\|>R\}\big)>\delta\big]\to 0\qquad\mbox{ as }R\to\infty.
\]
\end{lemma}
\begin{proof}[Sketch]
To avoid the idea being buried in notation, we sketch the proof in the 
special case of $d=1$ and $\DG=\Delta$.
For any $f\in C_b^\infty(\IR)$,
\begin{multline}
\label{special case first moment equation}
\IE[\langle f,\eta_t\rangle] 
=\IE[\langle f,\eta_0\rangle]
\\
+\IE\Big[\int_0^t\Big\langle\gamma(x, \smooth{\gamma}\eta_s(x))
\Delta \big(f(\cdot)r(\cdot,\smooth{r}\eta_s(\cdot))\big)(x)
+F(x,\smooth{F}\eta_s(x)) f(x),
\eta_s(dx)\Big\rangle ds\Big].
\end{multline} 
We take a sequence of test functions $f_n$ increasing to
$\exp(\lambda x)$, with $\lambda>0$. Having passed
to the limit,
substituted in~(\ref{special case first moment equation}) and 
expanded the term involving the Laplacian, we obtain
\begin{multline}
\label{exponential moment}
    \IE[\langle \exp(\lambda x),\eta_t\rangle]
    =
    \IE[\langle \exp(\lambda x),\eta_0\rangle]
    +
    \IE\Big[\int_0^t\Big\langle\gamma(x, \smooth{\gamma}\eta_s(x))
    \exp(\lambda x)
    \Big\{
        \lambda^2 r(x, \smooth{r}\eta_s(x))
    \\ {}
        + 2\lambda\partial_x\big(r(\cdot,\smooth{r}\eta_s(\cdot)\big)(x)
        + \partial_{xx} r(\cdot,\smooth{r}\eta_s(\cdot)\big)(x)
    \Big\} 
    + F(x,\smooth{F}\eta_s(x)),
    \eta_s(dx)\Big\rangle ds\Big] .
\end{multline} 
    \comment{alison to look at this again; check whether it depends on just bounding $B_1$}
Evidently, in much the same way as in the proof of
Lemma~\ref{lem:eta_f_bound},
our conditions on $r$ and $F$ ensure
that the integrand on the right hand side 
of~(\ref{exponential moment})
is bounded by $C\langle\exp(\lambda x),\eta_s(dx)\rangle$ for some constant
$C$. Gronwall's inequality then implies that 
$\IE[\langle\exp(\lambda x),\eta_t(dx)\rangle]$ 
is uniformly bounded for $t\in [0,T]$, and an application of Markov's inequality
tells us that for any $\delta >0$,
$\IP[\eta_t([R,\infty])>\delta]\to 0$ as $R\to\infty$. A symmetric argument
yields the corresponding bound for $\IP[\eta_t([-\infty, -R])>\delta$ and the 
result is proved.
\end{proof}

@@

\comment{below $\eta_s^N$ stops having  an $N$ at some point}

@@

For the second claim, first define $\tau(t) = T(t)-t$.
Since if $\tau(t) \ge -(\epsilon^2 + t)/2$,
then $1/(\epsilon^2 + T(t)) \le 2/(\epsilon^2 + t)$,
while $\epsilon^2 + T(t) \ge \epsilon^2$ always,
then partitioning over
$\{ \tau(t) \ge -(\epsilon^2 + t)/2 \}$ and its complement,
\begin{align}
   \| \psi^{\epsilon,x} \|_\infty
    &=
        \IE\left[ 
            \frac{1}{\big(2 \pi (\epsilon^2+T(t))\big)^{d/2}}
        \right] \nonumber
    \\ & \le
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        + 
        \frac{C}{\epsilon^d}
        \IP\left\{
            \tau(t) < - (\epsilon^2 + t) / 2
        \right\} \label{eqn:psi_infty_bound}
\end{align}
Now, observe that since $\IE[e^{-\Pi(\theta t)}] = \exp(-\theta t (1 - e^{-1}))$,
by Markov's inequality,
\begin{align}
    \IP\left\{
        \tau(t) < - \frac{\epsilon^2 + t}{2}
    \right\}
    &=
    \IP\left\{
        e^{-\Pi(\theta t)} > e^{-\theta (t - \epsilon^2) / 2}
    \right\}
\nonumber
\\
&\leq \frac{\IE[\exp\big(-\Pi(\theta t)\big)]}{\exp\big(-\theta(t-\epsilon^2)/2\big)}
\nonumber
    \\&=
    \frac{
        \exp(- \theta t(1 - e^{-1}))
    }{
        \exp(- \theta (t - \epsilon^2) / 2)
    }
\nonumber
    \\&=
    \exp\left\{ - \delta \theta t - \frac{\theta \epsilon^2}{2} \right\} ,
\label{bound for negative tau}
\end{align}
where $\delta = 1/2 - e^{-1} > 0$.
The second term in \eqref{eqn:psi_infty_bound} is therefore bounded by
\[
    C \left(1 + \frac{t}{\epsilon^2}\right)^{d/2} e^{-\delta \theta t - \epsilon^2 \theta / 2}
    \frac{1}{(\epsilon^2 + t)^{d/2}} .
\]
Now observe that the derivative (with respect to $t$) of 
$e^{-\delta \theta t - \epsilon^2 \theta / 2} (1 + t/\epsilon^2)^{d/2}$ 
is
\[
    \left( \frac{d}{2 \epsilon^2} - \left(1 + \frac{t}{\epsilon^2}\right) \delta \theta \right)
    \left( 1 + \frac{t}{\epsilon^2} \right)^{d/2 - 1} e^{-\delta \theta t - \epsilon^2 \theta / 2} ,
\]
which is negative if $\theta(\epsilon^2 + t) > d/2 \delta$.
At the maximum, $(\epsilon^2 + t) = d \delta / (2 \theta)$,
so $(1 + t/\epsilon^2) = d \delta/(2 \theta \epsilon^2)$,
and so this quantity is bounded uniformly over not only $t$ but also $\epsilon$.
\comment{explain that's where we use bounded-below-ness of $\theta \epsilon^2$.}
Therefore, the second term in \eqref{eqn:psi_infty_bound} 
is bounded by $C / (\epsilon^2 + t)^{d/2}$
(for a different constant $C$).
Substituting this into \eqref{eqn:psi_infty_bound} yields the result.
\end{proof}

@@

\comment{Probably should put a bit more detail in here?}
\comment{key bit is that we can always write $M_s(w)$ as a convolution of $p_{\epsilon^2/}$ with something}
